---
title: "Lab 1 - Exploring Data"
author: "Vipul Mehra"
date: "9/19/2017"
output:
  html_document: default
  pdf_document: default
---
## Include Paths
Setting the working directory

```{r}
work_dir <- "~/Documents/ISTE-780/Labs/Lab-1/lab1/"
setwd(work_dir)
```

## Loading Data
Reading data from CSV file. Saving it to variable named 'college'.

```{r}
college <- read.csv("/Users/vipulmehra/Documents/ISTE-780/Labs/Lab-1/lab1/College.csv")
```

## Summary of the dataset
Going through summary of the dataset

```{r}
summary(college)
```

## Trying fix and adding a rownames column with name of each university recorded

```{r}
rownames(college)=college[,1]
fix(college)
```

## Eliminating first column from the dataset where names are stored

```{r}
college=college[,-1]
fix(college)
```

## Getting numerical summary of the variables in dataset

```{r}
summary(college)
```

## Getting scatterplot of the first 10 variables of the data

```{r}
pairs(college[,1:10])
```

## Plotting box plot of Outstate vs Private

```{r}
boxplot(college$Outstate ~ college$Private, col = c("Grey", "Red"), main = "Outstate versus Private", xlab = "Private", ylab = "Outstate")
```

## Creating qualitative variable namely Elite by binning Top10perc variable

```{r}
Elite <-  rep("No",nrow(college))
Elite[college$Top10perc > 50] <-  "Yes"
Elite <- as.factor(Elite)
college <- data.frame(college, Elite)
fix(college)
```

## Using the summary() function to see how many elite universities there are. And plotting boxplot to produce side-by-side boxplots of Outstate versus Elite.
Printing summary

```{r}
summary(college$Elite)
```

Printing the boxplot of Outstate versus Elite

```{r}
boxplot(college$Outstate ~ college$Elite, col = c("Grey", "Red"), main = "Outstate versus Elite", xlab = "Elite", ylab = "Outstate")
```

## Printing the Histograms of few of the quantitative variables with different values of break or with different binning

```{r}
#Dividing the window
par(mfcol = c(2,6))

#Quantitative Variables with 5 bins
hist(college$Accept, breaks = 5, freq = TRUE, col = "Grey", main = "Histogram Bins of 5", 
    xlab = "Accept", ylab = "Value")
hist(college$Enroll, breaks = 5, freq = TRUE, col = "Grey", main = "Histogram Bins of 5", 
    xlab = "Enroll", ylab = "Value")
hist(college$Top10perc, breaks = 5, freq = TRUE, col = "Grey", main = "Histogram Bins of 5", 
    xlab = "Top10perc", ylab = "Value")
hist(college$Top25perc, breaks = 5, freq = TRUE, col = "Grey", main = "Histogram Bins of 5", 
    xlab = "Top25perc", ylab = "Value")

#Quantitative Variables with 10 bins
hist(college$Accept, breaks = 10, freq = TRUE, col = "Red", main = "Histogram Bins of 10", 
    xlab = "Accept", ylab = "Value")
hist(college$Enroll, breaks = 10, freq = TRUE, col = "Red", main = "Histogram Bins of 10", 
    xlab = "Enroll", ylab = "Value")
hist(college$Top10perc, breaks = 10, freq = TRUE, col = "Red", main = "Histogram Bins of 10", 
    xlab = "Top10perc", ylab = "Value")
hist(college$Top25perc, breaks = 10, freq = TRUE, col = "Red", main = "Histogram Bins of 10", 
    xlab = "Top25perc", ylab = "Value")

#Quantitative Variables with 15 bins
hist(college$Accept, breaks = 15, freq = TRUE, col = "Black", border = "White", main = "Histogram Bins of 15", 
    xlab = "Accept", ylab = "Value")
hist(college$Enroll, breaks = 15, freq = TRUE, col = "Black", border = "White", main = "Histogram Bins of 15", 
    xlab = "Enroll", ylab = "Value")
hist(college$Top10perc, breaks = 15, freq = TRUE, col = "Black", border = "White", main = "Histogram Bins of 15", 
    xlab = "Top10perc", ylab = "Value")
hist(college$Top25perc, breaks = 15, freq = TRUE, col = "Black", border = "White", main = "Histogram Bins of 15", 
    xlab = "Top25perc", ylab = "Value")
```

## Summary to find qualitative (categorical) and quantitative predictors

```{r}
summary(college)
```

As can be seen from the above summary Elite and Private are the 2 Predictors that are categorical or we can say qualitative and have the options Yes and No. All other which includes Apps, Accept, Enroll, Top10perc, Top25perce, F.Undergrad, P.Undergrad, Outstate, Room.Board, Books, Personal, PhD, Terminal, S.F.Ration, perc.alumni, Expend, Grad.Rate are quantitative. 

Please note: Originally, we had only 1 qualitative that is Private. But, after binning Top10perc we created another qualitative variable that is Elite.

## Finding Range, Mean and Standard deviation of Enroll

```{r}
#Range of Enroll: Finds the min and max value and stores it in a numeric vector
range <- range(college$Enroll)
#We will take the difference of the max and the min value to get the range
diff(range)
#Find mean of Enroll
mean(college$Enroll)
#Find standard deviation of Enroll
sd(college$Enroll)
```

## Removing rows and finding Range, Mean, Standard deviation
Removing 100th through 200th observations from the college dataframe and storing in a new dataframe (so that we don't loose the observations in original dataframe) and then again finding Range, Mean and Standard deviation of Enroll in the subset of data that remains.

```{r}
#Removing 100th to 200th observations from the college dataframe and storing it as a college_new_df dataframe
college_new_df <- college[-(100:200), ]
#Range of Enroll: Finds the min and max value and stores it in a numeric vector
range_new_df <- range(college_new_df$Enroll)
#We will take the difference of the max and the min value to get the range
diff(range_new_df)
#Find mean of Enroll in the college_new_df dataframe
mean(college_new_df$Enroll)
#Find standard deviation of Enroll in the college_new_df dataframe
sd(college_new_df$Enroll)
```

## Alternate method of finding Range, Mean, Standard deviation
Here we are not creating a new dataframe

```{r eval=FALSE, echo=T}
#Removing 100th to 200th observations from the college dataframe and storing it as a college_new_df dataframe
college[-(100:200), ]
fix(college)
#Range of Enroll: Finds the min and max value and stores it in a numeric vector
range <- range(college$Enroll)
#We will take the difference of the max and the min value to get the range
diff(range)
#Find mean of Enroll in the college_new_df dataframe
mean(college$Enroll)
#Find standard deviation of Enroll in the college_new_df dataframe
sd(college$Enroll)
```

